---
title: "report2"
author: "Nathaniel Brown, Huijia Yu, Angie Shen"
date: "November 6, 2017"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message=FALSE)
library(reshape2)
library(ggplot2)
library(dplyr)
library(mice)
library(micemd)
library(VIM)
library(knitr)
```

# Introduction

We are interested in predicting whether malaria parasites will be found in a child's blood based on his/her age, bednet use, the amount of greenery in his/her village, and the presence of a health clinic. However, almost 40% of the observations do not have bednet use reported, so removing those observations would cause us to lose too much information. Instead, we will impute bednet values to predict the outcome.


```{r}
dat <- read.csv("gambiaMissing.csv")
dat$Y = factor(dat$Y)
dat$BEDNET = factor(dat$BEDNET)
dat$PHC = factor(dat$PHC)
dat$GREEN = factor(as.character(formatC(dat$GREEN, format='f', digits=2)))

```


# Exploratory Data Analysis

In the graphs below, we illustrate the frequency of each category within each the dataset:

```{r, fig.height=2, fig.width=3}
ylabel <- dat %>% group_by(Y) %>% summarize(n=n()) %>% '[['(2) %>% as.numeric()
agelabel <- dat %>% group_by(AGE) %>% summarize(n=n()) %>% '[['(2) %>% as.numeric()
bednetlabel <- dat %>% group_by(BEDNET) %>% summarize(n=n()) %>% '[['(2) %>% as.numeric()
greenlabel <- dat %>% group_by(GREEN) %>% summarize(n=n()) %>% '[['(2) %>% as.numeric()
phclabel <- dat %>% group_by(PHC) %>% summarize(n=n()) %>% '[['(2) %>% as.numeric()

ggplot(data=dat, aes(x=Y)) + 
  geom_bar(fill = "gold") + 
  geom_text(data = data.frame(), aes(x=sort(unique(dat$Y)), y=ylabel/2, label=ylabel)) 

ggplot(data=dat, aes(x=AGE)) + 
  geom_bar(fill = "red") + 
  geom_text(data = data.frame(), aes(x=sort(unique(dat$AGE)), y=agelabel/2, label=agelabel))

ggplot(data=dat, aes(x=BEDNET)) + 
  geom_bar(fill = "blue") + 
  geom_text(data = data.frame(), aes(x=sort(unique(dat$BEDNET), na.last=TRUE), y=bednetlabel/2, label=bednetlabel))

ggplot(data=dat, aes(x=GREEN)) + 
  geom_bar(fill = "green") + 
  geom_text(data = data.frame(), aes(x=sort(unique(dat$GREEN)), y=greenlabel/2, label=greenlabel))

ggplot(data=dat, aes(x=PHC)) + 
  geom_bar(fill = "purple") + 
  geom_text(data = data.frame(), aes(x=sort(unique(dat$PHC)), y=phclabel/2, label=phclabel))

```


Below, we investigate the proportions of missing bednet responses by each level of the predictors.

```{r, fig.height=2, fig.width=3}
dat <- read.csv("gambiaMissing.csv")

bednet_group <- function(dat= NA, predictor = NA){
  bednet_group <- dat %>% 
                group_by_(predictor) %>% 
                summarize(#BEDNET0 = mean(BEDNET==0 & !is.na(BEDNET)), 
                          #BEDNET1 = mean(BEDNET==1 & !is.na(BEDNET)), 
                          BEDNET01 = mean(!is.na(BEDNET)),
                          BEDNETNA = mean(is.na(BEDNET))) %>%
                melt(id=predictor)
  colnames(bednet_group) <- c("x", "missing", "y")
  return(bednet_group)
}

bednet_age <- bednet_group(dat, "AGE")
bednet_green <- bednet_group(dat, "GREEN")
bednet_phc <- bednet_group(dat, "PHC")
bednet_y <- bednet_group(dat, "Y")

ggplot(data=bednet_age, aes(x=as.factor(x), y=y, fill=missing)) + 
  geom_col(position="stack") + 
  labs(title="Age by Bednet",x="Age", y="Proportion", fill = "Missing") + 
  scale_fill_discrete(labels = c("not missing", "missing")) +
  theme(plot.title = element_text(hjust = 0.5))

ggplot(data=bednet_green, aes(x=as.factor(x), y=y, fill=missing)) +
  geom_col(position="stack") + 
  labs(title="Greenery by Bednet",x="Green", y="Proportion", fill = "Missing") + 
  scale_fill_discrete(labels = c("not missing", "missing")) +
  theme(plot.title = element_text(hjust = 0.5), axis.text.x=element_text(size=5))

ggplot(data=bednet_phc, aes(x=as.factor(x), y=y, fill=missing)) + 
  geom_col(position="stack") +   
  labs(title=" Health Clinic by Bednet",x="PHC", y="Proportion", fill = "Missing") + 
  scale_fill_discrete(labels = c("not missing", "missing")) +
  theme(plot.title = element_text(hjust = 0.5))

ggplot(data=bednet_y, aes(x=as.factor(x), y=y, fill=missing)) + 
  geom_col(position="stack") +   
  labs(title="Parasites by Bednet",x="Malaria", y="Proportion", fill = "Missing") + 
  scale_fill_discrete(labels = c("not missing", "missing")) +
  theme(plot.title = element_text(hjust = 0.5))

```

# Multivariate Imputation by Chained Equations (MICE)

```{r}
m <- 10
temp <- mice(dat, m = m, maxit = 35,method = c("", "", "logreg", "", ""), print = FALSE)
#summary(temp)
plot(temp)
```

```{r, fig.height=3,fig.width=5}

densityplot(temp)

```

The trace lines appear to be stationary and free of trends, indicating convergence. We can see that the density distribution of each of the imputed datasets (in red) is congruent with the original one (in blue). 

# Predicting Malaria

To predict whether someoneone will get malaria and to calculate the effect of each predictor, we use a logistic regression model, which is applied to each generated dataset individually.

```{r}
model0 = with(temp, glm(Y ~ 1, family="binomial"))
model1sub = with(temp, glm(Y ~ as.factor(AGE) + as.factor(GREEN) + as.factor(PHC), family="binomial"))
model1 = with(temp, glm(Y ~ as.factor(AGE) + as.factor(GREEN) + as.factor(PHC) + as.factor(BEDNET), family="binomial"))

resids <- sapply(model1[[4]], residuals)
fitvals <- sapply(model1[[4]], fitted.values)


par(mfrow = c(4,3))
for(i in 1:m){
  arm::binnedplot(fitvals[,i], resids[,i], main = paste0("Binned Residuals (m=",i,")"))
}

pool_coef <- as.matrix(summary(pool(model1))[,"est"])
dat_complete <- dat[complete.cases(dat),]
dat_complete_matrix <- (as.matrix(model.matrix(Y ~ as.factor(AGE) + as.factor(GREEN) + as.factor(PHC) + as.factor(BEDNET), data=dat_complete)))
fitvals <- arm::invlogit(t(as.matrix(pool_coef)) %*% t(dat_complete_matrix))
resids <- as.numeric(dat_complete[,1]) - fitvals - 1

  arm::binnedplot(as.numeric(fitvals), as.numeric(resids), main = paste0("Pooled Binned Residuals"))

```

Neither the pooled nor the individual residuals display any trends or changes in variance, which means this model is appropriate.

We can also check the accuracy of the model by looking at the confusion matrix for the pooled model.

```{r}
confusion <- table(dat_complete$Y, round(fitvals))
fourfoldplot(confusion)
```

The model is correct most of the time, but it has a lot of false negatives.

```{r}
plot(pROC::roc(dat_complete$Y, as.numeric(fitvals)),
     main = "ROC Curve for Malaria Predictions")
```

The ROC curve shows us that the predictions are not too accurate.


# Inference on Coefficients

```{r}
kable(summary(pool(model1)) %>% '['(,-c(ncol(.), ncol(.)-1)))

```


To further confirm that the bednet variable is not significant but the others are, we use a likelihood test comparing the null model to the one without BEDNET, and then one comparing models with and without BEDNET.

```{r}
pool.compare(model1, model1sub, data=temp, method="likelihood")$pvalue
pool.compare(model1sub,model0, data=temp, method="likelihood")$pvalue

```

After fitting a logistic regression to each of the 10 generated datasets, we can see that the bednet variable is actually not significant at the $\alpha$=0.05 confidence level.

# Contributions

Nathaniel made the Exploratory Data Analysis plots on the missingness of the bednet variable, and wrote the observations. Huijia worked on the Discussion of Approaches section. Angie worked on chained regression.

# References

Azur, M. J., Stuart, E. A., Frangakis, C., & Leaf, P. J. (2011). Multiple Imputation by Chained Equations: What is it and how does it work? International Journal of Methods in Psychiatric Research, 20(1), 40–49. http://doi.org/10.1002/mpr.329
